### 2020-09-24 paper/figures/fq_common_end.tex
Updated paper

%\draw[draw,decorate,decoration={brace,mirror}] (0.75,0) -- (1.75,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm,font=\fontsize{3pt}{0}\selectfont] {interval};
%\draw[densely dotted] (1.75,0) -- (1.75,0.95);

\newcommand{\xaxislength}{1.8}

\draw[->, line width=0.05mm, -dot3-={0.5 cm}{1 rtt}, -dot3-={1.0 cm}{2 rtts}, -dot3-={1.5 cm}{3 rtts}] (0,0) -- ++(\xaxislength , 0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1.65, 0.05) (time) {time};

\draw[->, line width=0.05mm] (0,0) -- ++(0,2.35);
\draw[draw=none, line width=0.05mm] (-0.05,-0.0) -- ++(0,1);
\draw[draw=none, line width=0.05mm] (0,-0.05) -- ++(0,1);

\node (red_label) at (0.1,1.95) [inner sep=0,fill=red,rectangle,minimum width=0.1cm,minimum height=0.1cm,label={[align=left,label distance=-0.75mm]right:\fontsize{1mm}{0mm}\selectfont{}flow 2}] {};

\node (green_label) at (0.1,1.8) [inner sep=0,fill=green,rectangle,minimum width=0.1cm,minimum height=0.1cm,label={[align=left,label distance=-0.75mm]right:\fontsize{1mm}{0mm}\selectfont{}flow 1}] {};

\end{tikzpicture}
\end{document}